---
title: "The TwoRegression Package"
author: "Paul R. Hibbing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The TwoRegression Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## INTRODUCTION

The TwoRegression package allows users to quickly and accurately develop/apply
two-regression algorithms to data from research-grade wearable devices. This
vignette is designed to demonstrate  usage of the package's core features.

Before getting into that, it's valuable to cover some history. The package was
initially established as a home for the models of [Hibbing et al.
(2018)](https://www.ncbi.nlm.nih.gov/pubmed/29271847). Since the initial
release, support has been added for developing/applying new models, as well as
applying others from prior research (see [Crouter et
al.(2006)](https://pubmed.ncbi.nlm.nih.gov/16322367/), [Crouter et
al.(2010)](https://pubmed.ncbi.nlm.nih.gov/20400882/), and [Crouter et
al.(2012)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3324667/)). As of
version 1.0.0, a new approach has been implemented for invoking prior methods
via the `TwoRegression` function, which we will look at in the following
section. Afterwards, we will cover the process of creating and cross-validating
new models, plus other aspects of using them effectively.


## IMPLEMENTING TWO-REGRESSION MODELS THAT ALREADY EXIST

#### Approach and Options

Prior models are implemented using the `TwoRegression` function. Currently,
support is available for the following:

* [Crouter et al.(2006)](https://pubmed.ncbi.nlm.nih.gov/16322367/)
  - The original Crouter two-regression model (for adults)
* [Crouter et al.(2010)](https://pubmed.ncbi.nlm.nih.gov/20400882/)
  - The refined Crouter two-regression model (for adults)
* [Crouter et al.(2012)](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3324667/))
  - Two youth-specific Crouter two-regression models, one for vertical axis
    data and the other for vector magnitude data
* [Hibbing et al. (2018)](https://pubmed.ncbi.nlm.nih.gov/29271847/)
  - Fifteen two-regression models for adults, corresponding to one of five
    attachment sites (hip, wrists, ankles) and one of three model
    configurations (accelerometer only, accelerometer and gyroscope, or
    accelerometer, gyroscope, and magnetometer)
    
#### Help documentation

It's very important that you look at the `TwoRegression` function documentation.
It will help you understand what settings you need to provide in order to run a
specific model correctly. To view the documentation, run the following:

```{r, eval = FALSE}

?TwoRegression::TwoRegression

```

This will pull up a documentation page where you can see the syntax for calling
the `TwoRegression` function. Importantly, the page also lists the syntax for
several internal functions that apply to specific models (i.e., the
`crouter_2006`, `crouter_2010`, `crouter_2012`, and `hibbing_2018` functions).
The `TwoRegression` function is just a wrapper around those other internal
functions. That is, based on the method you select, `TwoRegression` will call
the corresponding internal function. In most cases, you will also need to
provide settings for the internal function, and those can be passed into the
`TwoRegression` function directly. This will be easier to see and understand in
the coding samples later on in this section, but it's important to be aware of
all this from the get-go.

#### Assumptions and Data

The function operates under the assumption you already have data read into R.
You can do this with the [AGread package](https://github.com/paulhibbing/AGread),
although installation of that package might be tricky. Feel free to [file an issue
on the GitHub webpage](https://github.com/paulhibbing/AGread/issues/new) if you need extra help.

Anyway, the TwoRegression package provides some sample data to show how you
might apply the models. If you can get your own data into a form that mirrors
the sample data, you'll be in good shape. Here's how to access these built-in
datasets:

```{r}

data(count_data, package = "TwoRegression")
data(all_data, package = "TwoRegression")

```

Notably, the `count_data` object contains activity count data (for the Crouter
two-regression models), while the `all_data` object contains raw sensor data
(for the Hibbing models). We can view the first few rows of count data as
follows:

```{r}

utils::head(count_data)

```

We can do the same using a similar approach for the raw data. However, let's
remove some extraneous variables as well, just for the sake of this
illustration.

```{r}

all_data <- all_data[ ,setdiff(
  names(all_data),
  ## These are the variables to remove:
  c(
    "file_source_PrimaryAccel", "date_processed_PrimaryAccel",
    "file_source_IMU", "date_processed_IMU", "day_of_year", "minute_of_day",
    "ENMO_CV10s", "GVM_CV10s", "Direction"
  )
)]

utils::head(all_data)
```

#### Applying Crouter models (for activity count data)

Once you have your dataset ready, it's easy to apply a two-regression model.
Just invoke the `TwoRegression` function like this:

```{r}

crouter2006 <- TwoRegression::TwoRegression(
  count_data, "Crouter 2006", movement_var = "Axis1", time_var = "time"
)

crouter2010 <- TwoRegression::TwoRegression(
  count_data, "Crouter 2010", movement_var = "Axis1", time_var = "time"
)

crouter2012_va <- TwoRegression::TwoRegression(
  count_data, "Crouter 2012", movement_var = "Axis1",
  time_var = "time", model = "VA", check = FALSE
)

crouter2012_vm <- TwoRegression::TwoRegression(
  count_data, "Crouter 2012", movement_var = "Vector.Magnitude",
  time_var = "time", model = "VM", check = FALSE
)

```

For the Crouter 2012 models, you have to choose between the vertical axis model
and the vector magnitude model. If you don't set `check = FALSE`, you will get a
warning about which movement variable and model you've selected. This is meant
as a prompt for you to ensure your selected movement variable matches your
selected model. Once you're confident in your selection, you can set `check =
FALSE` and the warning won't show up.

For the time being, you can only implement Crouter models one at a time. Of
course, you can combine the output from multiple models yourself. Ideally, with
ongoing development, a point will come where this can be done automatically and
efficiently (see the [GitHub issue on this topic](https://github.com/paulhibbing/TwoRegression/issues/1)), but for now it
isn't built in. As we'll see in the following subsection, though, that's not the
case for the Hibbing models. Here's a look at the output from the prior commands:

```{r}

utils::head(crouter2006)
utils::head(crouter2010)
utils::head(crouter2012_va)
utils::head(crouter2012_vm)

```

#### Applying Hibbing models (for raw sensor data)

The Hibbing models are implemented similarly to the Crouter models. A key
difference, though, is that you can ask the function to run multiple models
simultaneously. That's what we'll see in the following example:

```{r}

hibbing2018 <- TwoRegression::TwoRegression(
  
  all_data, "Hibbing 2018", accel_var = "ENMO",
  gyro_var = "Gyroscope_VM_DegPerS",
  direction_var = "mean_magnetometer_direction",
  
  ## Here is where we can select an algorithm from multiple sites:
  site = c("Left Ankle", "Right Ankle"),
  
  ## And here is where we can select multiple algorithms
  ## (1 = accelerometer only; 2 = accelerometer and gyroscope;
  ## 3 = accelerometer, gyroscope, and magnetometer)
  algorithm = 1:2,
  
  ## We can also ask the function to collapse data every minute by making an
  ## extra call to `smooth_2rm`
  smooth = TRUE
)

utils::head(hibbing2018)

```

So, each algorithm is run, and the information is stored in a unique and
descriptive variable name.

## FITTING AND EXAMINING NEW MODELS

#### Background and Setup

The TwoRegression package is also useful if you want to create your own model.
Your dataset needs to have some more complex information to get this going,
though. Using our previous `all_data` object, we need to label it with some
pretend activity labels and some pretent energy expenditure values in metabolic
equivalents (METs). In a real-life setting, the MET values would likely come
from indirect calorimetry. To create some of this imaginary data, we can run the following:

```{r}

set.seed(307)

fake_sed <- c("Lying", "Sitting")
fake_lpa <- c("Sweeping", "Dusting")
fake_cwr <- c("Walking", "Running")
fake_ila <- c("Tennis", "Basketball")

fake_activities <- c(fake_sed, fake_lpa, fake_cwr, fake_ila)

all_data$Activity <- sample(fake_activities, nrow(all_data), TRUE)

all_data$fake_METs <- ifelse(
  all_data$Activity %in% c(fake_sed, fake_lpa),
  runif(nrow(all_data), 1, 2),
  runif(nrow(all_data), 2.5, 8)
)

```

There are a couple of other steps we need to take care of that are somewhat
specific to this demonstration. Nevertheless, they may be helpful to see. One is
to make sure our data have a column indicating which participant each data point
came from. In this case, we'll just label our data to pretend it came from two
sample files instead of one (where 'sample file' is analogous to 'participant').
The other step is calculating the coefficient of variation. We technically
deleted that variable earler, but here we'll see how to add it back in.

```{r}

all_data$PID <- rep(
  c("Test1", "Test2"),
  each = ceiling(nrow(all_data) / 2)
)[seq(nrow(all_data))]

all_data$ENMO_CV10s <- TwoRegression::cv_2rm(all_data$ENMO)

```

#### Fitting the Model

When we go to fit the model, we'll use the `fit_2rm` function. There are a lot
of arguments to provide here:

* **data: ** The dataset
* **activity_var: ** The name of the variable that indicates the activity
* **sed_cp_activities: ** The subset of values from `activity_var` that should
  be included when calibrating the 2RM sedentary cut point
* **sed_activities: ** The subset of values from `activity_var` that should be
  labeled as positive for sedentary behavior when calibrating the 2RM sedentary
  cut point
* **sed_cp_var: ** The name of the variable for which the 2RM sedentary cut
  point should be calibrated.
* **sed_METs: ** The MET value to assign for sedentary behaviors (i.e., when
  the value for `sed_cp_var` falls below the 2RM sedentary cut point)
* **walkrun_activities: ** The subset of values from `activity_var` that should
  be labeled as positive for "continuous walking/running" (CWR) when
  calibrating the 2RM CWR cut point
* **walkrun_cp_var: ** The name of the variable for which the 2RM CWR cut point
  should be calibrated
* **met_var: ** The name of the MET variable that the 2RM should be fitted
  to predict
* **walkrun_formula: ** A character representation of the formula that should
  be used when fitting the CWR model (formulated as `outcome ~ predictors`)
* **intermittent_formula: ** A character representation of the formulat that
  should be used when fitting the intermittent lifestyle activities model
  (formulated like `walkrun_formula` -- note that data transformations like
  squaring or cubing should be wrapped in `I()`)
  
From there, we can fit our model like this:

```{r}

my_model <- TwoRegression::fit_2rm(
  data = all_data,
  activity_var = "Activity",
  sed_cp_activities = c(fake_sed, fake_lpa),
  sed_activities = fake_sed,
  sed_cp_var = "ENMO",
  sed_METs = 1.25,
  walkrun_activities = fake_cwr,
  walkrun_cp_var = "ENMO_CV10s",
  met_var = "fake_METs",
  walkrun_formula = "fake_METs ~ ENMO",
  intermittent_formula = "fake_METs ~ ENMO + I(ENMO^2) + I(ENMO^3)"
)

```

#### Examining Model Performance

The package provides summary and plot methods to understand, cross-validate, and
visualize the model. Notably, this demonstration model is not meant to perform
well or look pretty (the data are just numbers that have no real meaning), but
we'll still take a look at how to run the code.

As far as the summary method goes, this is where we need the participant
identification column we set up earlier. Specifically, it will be used for
leave-one-out cross-validation, where the data are split up into different
chunks while the model is repeatedly re-fitted. Other information in the output
includes a textual representation of the overall algorithm and summaries of the
fit/performance of individual components (i.e., ROC and regression analyses). To
pull all of this up, you just have to run code that matches the following
pattern:

```{r}

summary(
  my_model,
  subject_var = "PID",
  MET_var = "fake_METs",
  activity_var = "Activity"
)

```

For the plot function, you'll need to fill in some of the same values from the
original call to `fit_2rm`. Use code that matches the following pattern:

```{r, fig.align='center', fig.width=12, fig.height=6, out.width="85%"}

## You have to explicitly type `object = ` for this to work
plot(
  object = my_model,
  sed_cp_activities = c(fake_sed, fake_lpa),
  sed_activities = fake_sed,
  sed_cpVar = "ENMO",
  activity_var = "Activity",
  met_var = "fake_METs",
  walkrun_activities = fake_cwr,
  walkrun_cpVar = "ENMO_CV10s",
  print = TRUE
)

```

## USING NEW MODELS

Once you've created your model, you want to use it on new data. That's easy to
do using the `predict` method included in the package. If we pretend our
`all_data` object is a new dataset, we could get predictions by running code
like this:

```{r}

results <- predict(my_model, all_data)

utils::head(results)

```

When making predictions, you can specify `verbose = TRUE` if you want to print a
message to the console about making predictions from your model. By default, it
will say it's making predictions using the 'user_unspecified' model. To give
your model a name, you can assign a value to its `method` element. Consider the following:

```{r}

results_default <- predict(my_model, all_data, verbose = TRUE)


my_model$method <- "My Customized 2RM"


results_updated <- predict(my_model, all_data, verbose = TRUE)

```

## CONCLUSION

That's it. This has been a quick crash course in the core features and functions
of the TwoRegression package. If you have questions or feedback, feel free to
connect by [posting an issue on the GitHub page](https://github.com/paulhibbing/TwoRegression/issues/new). Happy coding!
